---
title: "Statistical inference project. Part 1."
author: "Anton"
date: "Wednesday, September 03, 2014"
output:
  pdf_document: default
  html_document: default
---
#Simulation exercises  
The exponential distribution simulated with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is  $1/\lambda$ and the standard deviation is also $1/\lambda$. For all of the simulations $\lambda$ set to 0.2. 
Through 1000 simulation of averages of 40 exponential we try to answer four questions:  

```{r echo = FALSE}
lambda = 0.2
n = 40
nsim = 1000
tmn <- 1/lambda
tsd <- (1/lambda)/sqrt(n)
tvar <- tsd^2
set.seed(5639)
sim <- matrix(rexp(nsim*n, lambda), ncol = n) 
exp_mns <- rowMeans(sim)
mn <- mean(exp_mns)
sd <- sd(exp_mns)
var <- var(exp_mns)
```  

####1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.    

The distribution of sample means is centered at `r mn` while the theoretical center of the distribution is `r tmn`. So it's very close to the theoretical center of the distribution.  
   

####2. Show how variable the distribution and compare it to the theoretical variance of the distribution.  

The sample variance is `r var` while the theoretical variance is `r tvar`. Also standart deviation is `r sd` for the sample distribution and `r tsd` for the theoretical distribution.  

####3. Show that the distribution is approximately normal.  
Compare our simulated distribution and standart normal distribution by plotting scaled simulated means.  
Also compare two distributions (simulated and normal) by plotting their quantiles against each other.  
```{r echo = FALSE, fig.height = 4}
#require(ggplot2)
par(mfrow=c(1,2))
hist(exp_mns, freq = FALSE, col = "light blue", main = NULL, xlab = "Mean")
lines(density(exp_mns), col = "blue")
curve(dnorm(x, tmn, tsd), col="red", lty = 3, lwd = 2, add = TRUE)
qqnorm(exp_mns, main="", col = "blue")
qqline(exp_mns, col="red", lty = 3, lwd = 2)
```  

So on left figure we can see that our simulated distribution pretty close to normal distribution. The Q-Q plot (right figure) shows that most of the data points are on or near the straight line, suggests that the data is almost normally distributed.  

####4. Evaluate the coverage of the confidence interval for $1/\lambda = \bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$.    

```{r echo = FALSE}
ci95 <- mn + c(-1,1)*qnorm(0.975)*sqrt(var)/sqrt(length(exp_mns))
```  
   
The 95% confidence interval is `r ci95` for simulated distribution and the true mean is `r tmn`.  
Now look at the distribution of confidence intervals for each simulation against the true mean.  
```{r echo = FALSE, fig.height = 4}
ci <- data.frame("low" = NULL,'high' = NULL)
for(i in 1:nsim){
    ci[i,"low"]<- mean(sim[i,])-1.96*sd(sim[i,])/sqrt(n)
    ci[i,"high"]<- mean(sim[i,])+1.96*sd(sim[i,])/sqrt(n)
    }

coverage <- mean(ci$low < tmn & tmn < ci$high)*100
plot(ci, col = "blue")
abline(h = tmn, col="red", lty = 3, lwd = 2)
abline(v = tmn, col="red", lty = 3, lwd = 2)
```  

The coverage of all simulated confidence intervals is only `r coverage`%.  

------
\pagebreak  


#Appendix  
#### Full R code.  

```{r echo=TRUE, results='hide', fig.show='hide'}
lambda = 0.2
n = 40
nsim = 1000
tmn <- 1/lambda
tsd <- (1/lambda)/sqrt(n)
tvar <- tsd^2

set.seed(5639)
sim <- matrix(rexp(nsim*n, lambda), ncol = n) 
exp_mns <- rowMeans(sim)
mn <- mean(exp_mns)
sd <- sd(exp_mns)
var <- var(exp_mns)

#require(ggplot2)
par(mfrow=c(1,2))
hist(exp_mns, freq = FALSE, col = "light blue", main = NULL, xlab = "Mean")
lines(density(exp_mns), col = "blue")
curve(dnorm(x, tmn, tsd), col="red", lty = 3, lwd = 2, add = TRUE)
qqnorm(exp_mns, main="", col = "blue")
qqline(exp_mns, col="red", lty = 3, lwd = 2)

ci95 <- mn + c(-1,1)*qnorm(0.975)*sqrt(var)/sqrt(length(exp_mns))

ci <- data.frame("low" = NULL,'high' = NULL)
for(i in 1:nsim){
    ci[i,"low"]<- mean(sim[i,])-1.96*sd(sim[i,])/sqrt(n)
    ci[i,"high"]<- mean(sim[i,])+1.96*sd(sim[i,])/sqrt(n)
    }

coverage <- mean(ci$low < tmn & tmn < ci$high)*100
plot(ci, col = "blue")
abline(h = tmn, col="red", lty = 3, lwd = 2)
abline(v = tmn, col="red", lty = 3, lwd = 2)
```