---
title: "SIproject"
author: "Anton"
date: "Wednesday, September 03, 2014"
output: html_document
---
#Simulation exercises  
The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. Set lambda = 0.2 for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. Note that you will need to do a thousand or so simulated averages of 40 exponentials.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s.  You should
1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
2. Show how variable it is and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.
4. Evaluate the coverage of the confidence interval for 1/lambda: X¯±1.96Sn√.


```{r}
lambda = 0.2
n = 40
nsim = 1000
tmean <- 1/lambda
tsd <- (1/lambda)/sqrt(n)
tvar <- tsd^2
#rexp(n, lambda)
set.seed(45356)
sim <- matrix(rexp(nsim*n, lambda), ncol = n) 
#sim <- replicate(nsim,rexp(n,lambda))
exp_means <- rowMeans(sim)
mean <- mean(exp_means)
sd <- sd(exp_means)
var <- var(exp_means)
```  

####Show where the distribution is centered at and compare it to the theoretical center of the distribution.  
####Show how variable it is and compare it to the theoretical variance of the distribution.  

```{r}
mean
tmean
var
tvar
hist(exp_means, freq = FALSE, col = "light grey", border = "grey", main = NULL)
abline(v = mean, col = 'blue')
abline(v = tmean, col = 'red', lty = 3)
```  

####Show that the distribution is approximately normal. 
Two plots:  
compare our simulated distribution and standart normal distribution by plotting scaled simulated means  
comparing two probability distributions (simulated exponential and standart normal) by plotting their quantiles   against each other  
```{r}
nexp_means <- scale(exp_means)
par(mfrow=c(1,2))
hist(nexp_means, freq = FALSE, col = "light grey", border = "grey", main = NULL)
lines(density(nexp_means))
curve(dnorm(x), col='red', lty = 3, add = TRUE)
qqnorm(exp_means, main="")
qqline(exp_means)
```  

So we can see that our simulated distribution pretty close to normal distribution (not a standart because mean != 0 and standart deviation != 1).  

####Evaluate the coverage of the confidence interval for 1/lambda.  

```{r}
mean + c(-1,1)*qnorm(0.975)*sqrt(var)/sqrt(length(exp_means))
tmean + c(-1,1)*qnorm(0.975)*sqrt(tvar)/sqrt(nsim)
mean + c(-1,1)*qnorm(0.975)*sqrt(var)/sqrt(n)
tmean + c(-1,1)*qnorm(0.975)*sqrt(tvar)/sqrt(n)
```
The 95% confidence interval is .  
```{r}
ci <- data.frame("low" = NULL,'high' = NULL)
for(i in 1:nsim){
    ci[i,"low"]<- mean(sim[i,])-1.96*sd(sim[i,])/sqrt(n)
    ci[i,"high"]<- mean(sim[i,])+1.96*sd(sim[i,])/sqrt(n)
    }

coverage <- mean(ci$low < tmean & tmean < ci$high)
coverage
```